---
title: "Untitled"
author: "Ying-Jung Deweese"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)# trun off code echo
```

# Load packages

```{r,warning=F,message=F,echo=F}
library(ggplot2)
library(dplyr)
library(tidyr)
library(DT)
library(leaflet)
```

#Data Tidying

Read in data from Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M. 



```{r,echo=F,results=F}
esc<-read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e",method = "libcurl"),stringsAsFactors = F)

head(esc)

```

Calculate annual, regional total escapement by species

* separate the date into year -month -day
* 'group_by' 'summarize' to calculate the sum over unique region-year-species combinations
* 'filter' for salmon species

```{r}
annual_esc<-esc %>% 
  separate(sampleDate,into=c("year","month","day"),sep="-") %>% 
  mutate(year=as.numeric(year)) %>% 
  group_by(SASAP.Region,Species,year) %>% 
  summarize(escapment=sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook","Sockeye","Pink","Chum","Coho") ) # filter any species that are in






```


```{r,echo=F}
#echo=F hide code chunk
datatable(annual_esc)

```


# Static Plots


```{r}
my_theme<-  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5)) # this is the theme object
```



```{r}
ggplot(data=annual_esc,mapping = aes(x=Species,y=escapment,fill=SASAP.Region))+
  geom_col()+
  my_theme#fill=blue inside the geom column


```


time-series plot of Kodiak escapment by species

```{r}


ggplot(data=filter(annual_esc,SASAP.Region=="Kodiak"),
       aes(x=year,y=escapment,color=Species))+
  geom_line()+# fill for bar and column plot
  geom_point()+
  ylab("Escapement (number of fish)")+
  ggtitle("Kodiak Salmon Escapement")+
  my_theme

```

Plot timeseries of escapement by species for all regions


```{r,fig.height=10,fig.width=8,echo=F}
plot_name<-ggplot(annual_esc, aes(x = year, y = escapment, color = Species)) + 
    geom_line() +
    geom_point() +
    facet_wrap(~SASAP.Region, scales = "free_y",ncol=2) + #'free_y' make y axes different, facet_wrap (every plot inside is the same plot)
    ylab("Escapement") +
    my_theme

plot_name

ggsave("figures/region_escapment.png",
       plot=plot_name,
       height=10,
       width=8,
       units="in")

```

```{r}
write.csv(annual_esc,
          "derived_data/annual_escapement.csv",
          row.names = F)
```



# Interactive map

Make a map using leaflet

*find unique location with lat/lon values

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na() # drop rows contain NA for any column in your dataset

```


```{r}
datatable(locations)
```

Create our interactive map 

```{r,echo=F}
leaflet(locations) %>% #leaflet using pipe as the add sign in ggplot
  addTiles() %>% # add open source map tile in the background
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

